### 2013-08-21 Tex/chapters/acknowledgements.tex
Wortabstand hinter Dr. korrigiert

\addcontentsline{toc}{section}{Acknowledgements}
\section*{Acknowledgements}
    I would like to thank Christoph Norrenbrock and Dr.\ Oliver Melchert
    in whose office I worked on this thesis. Whenever I had a problem,
    it was no long way to ask one of them and their suggestions were
    always helpful. I am also grateful for the interesting subject, which
    was suggested by Christoph.\\
    Further I want to thank Prof.\ Dr.\ Alexander Hartmann, who offered the
    possibility to write this thesis in this working group,
    and the other members of this working group for the pleasant working
    atmosphere and conversations at lunch.\\
    This acknowledgement would not be complete if I had not thanked my
    parents. Whenever I told them about this thesis or showed some weird
    graphs, they were listening and truly interested.