### 2013-11-25 Tex/paper.tex
Umformulierungen, Korrketuren

\documentclass[prl,twocolumn,groupedaddress,showpacs,showkeys,amsmath,amssymb,floatfix]{revtex4-1}

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}

% \newcommand{\change}[1]{{\bf #1}}
% \newcommand{\changeAH}[1]{{\bf #1}}

\begin{document}

    \title{Critical Temperature of the Ising Ferromagnet on Proximity Graphs derived from Square Lattices by Site Displacement}
    \author{Hendrik Schawe}
    \email{hendrik.schawe@uni-oldenburg.de}
    \affiliation{Institut f\"ur Physik, Universit\"at Oldenburg, 26111 Oldenbug, Germany}
    \date{\today}

    \begin{abstract}
    \input{paper/abstract}
    \end{abstract}

    % TODO: recherchieren und ändern
    \pacs{64.60.ah}
    \maketitle

    \section{Introduction}
        \input{paper/introduction}

    \section{Model}
        \input{paper/model}

    \section{Methods}
        \input{paper/methods}

    \section{Results}
        \input{paper/results}

    \section{Conclusion}
        \input{paper/conclusion}

    \input{paper/acknowledgements}

    \bibliography{lit}
\end{document}