### 2013-12-19 Tex/paper/results.tex
Bilder kleiner gemacht

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=0.22\textwidth]{plots/RNG_Tc}
    \includegraphics[width=0.22\textwidth]{plots/GG_Tc}
    \caption[Critical Temperature over Different Disorder Parameters]
    {
        Critical temperatures \(T_c\) over different
        disorder parameters \(\sigma\) for
        Interesting points are the jump and rise of the GG at small
        \(\sigma\) and the plateau on the RNG for small \(\sigma\).
    }
    \label{fig:Tc}
\end{figure}

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=.45\textwidth]{plots/Binder_over_Correlation_All}
    \caption[Binder Cumulant over two-point finite-size correlation function divided by system size]
    {
        Binder Cumulant over two-point finite-size correlation function
        divided by system size. The datapoints for different system sizes
        are drawn with different symbols, while the datapoints for different
        graph ensembles are drawn with different colors.
        The inset zooms into the plot to show that
        all datapoints fall within errorbars onto one curve.
    }
    \label{fig:binderOverCorr}
\end{figure}