### 2015-08-31 Tex/paper/model.tex

\subsection{The Disordered Ising Model}
\label{ssec:model}
    We call the model under scrutiny in this article the \emph{disordered IFM}
    which is derived from the well understood \emph{square lattice IFM} \cite{Onsager1944}.
    That is a square lattice with lateral length \(L\) and \(N=L^2\)
    sites with periodic boundary conditions.
    Each site $i$ has an Ising spin \(s_i \in \{-1,+1\}\) and interacts with its
    nearest neighbors described by the Hamiltonian
    \begin{equation}
        \mathcal{H} = - \sum_{\avg{i,j}}J_{ij}s_{i}s_{j}.
        \label{eq:hamiltonian}
    \end{equation}
    \(\avg{i,j}\) refers to nodes \(i\) and \(j\) which are nearest
    neighbors, so that they are directly coupled to each other.
    This will be called an edge.
    The coupling between \(i\) and \(j\) is characterized by \(J_{ij} = J\),
    the coupling constant.

    The most important difference of the square lattice IFM to the
    disordered IFM is that the sites of the square lattice are
    displaced introducing geometric disorder resulting in a non regular graph
    structure.
    The displacement is randomly Gaussian distributed with standard
    deviation \(\sigma\), i.e.~the \(x\) and \(y\) coordinates of the
    sites are displaced by random \(\Delta x\) and \(\Delta y\) drawn
    from a Gaussian distribution.
    %~ \begin{equation}
        %~ f(\Delta x)=\frac{1}{\sqrt{2\pi}\sigma}\mathrm{e}^{-\frac{\Delta x^2}{2\sigma^2}}.
        %~ \label{eq:gauss}
    %~ \end{equation}
    %~ This is sketched in Fig.\ \ref{fig:displacement}.
    %~ \begin{figure}[htbp]
        %~ \centering
        %~ \includegraphics[width=0.45\textwidth]{images/Displacement}
        %~ \caption[Sketch how the Displacement works]
        %~ {
            %~ Sketch how the displacement of the nodes works. The nodes
            %~ get displaced by \(\Delta x\) and \(\Delta y\) drawn from the
            %~ distributions displayed next to the points. The original
            %~ square lattice is indicated by dashed lines.
        %~ }
        %~ \label{fig:displacement}
    %~ \end{figure}
    In the following \(\sigma\) will also be called \emph{disorder parameter}.
    If we take ``nearest neighbor" in the Euclidean meaning, most sites
    will only have one nearest neighbor after the
    displacement. If only the edges between these neighbors remained,
    the lattice would collapse to many very small clusters. But if one
    left the edges as they were before the displacement, the edges would
    cross -- at least for big displacements. The crossing of edges then
    would destroy the planar character of the model.
    To avoid this, a new edge set will be established after the displacement
    as shown in Fig.~\ref{fig:RNG_sigma} for example.
    \begin{figure}[htb]
        \centering
        \subfigure[$\sigma = 0.09$]
        {
            \label{sfig:RNG_sigma:0.09}
            \includegraphics[width=0.14\textwidth]{images/RNG/out009}
        }
        \subfigure[$\sigma = 0.15$]
        {
            \label{sfig:RNG_sigma:0.15}
            \includegraphics[width=0.14\textwidth]{images/RNG/out015}
        }
        \subfigure[$\sigma = 0.21$]
        {
            \label{sfig:RNG_sigma:0.21}
            \includegraphics[width=0.14\textwidth]{images/RNG/out021}
        }
        \caption[Examples of RNG for different $\sigma$]
        {
            Examples of proximity graphs (more precise, Relative Neighborhood
            graph see Sec. \ref{ssec:graphtypes}) for different $\sigma$.
            Connections which cross a periodic boundary are indicated
            by edges which connect a solid node to a dashed node.
        }
        \label{fig:RNG_sigma}
    \end{figure}
    The edges are constructed
    so that for a given configuration of points the resulting graph is an
    instance of a so called proximity graph, which will be explained in Sec.~\ref{ssec:graphtypes}.
    The coupling constant \(J_{ij}\) gets
    identified with edge weights. \(J_{ij}\) depends on the
    geometric distance between the connected pair of sites. More precise,
    the weight of an edge is
    \begin{equation}
        E_{ij} = J_{ij} = \mathrm{e}^{\alpha (1-d_{ij})}
        \label{eq:coupling}
    \end{equation}
    where \(d_{ij}\) is the Euclidean
    distance between the nodes \(i\) and \(j\). Following Ref.~\cite{Lima2000}
    the free parameter \(\alpha\) is set to \(\alpha = 0.5\).
    Note that for \(\sigma = 0\) all \(d_{ij} = 1\) and therefore all
    \(J_{ij} = 1\), so that the disordered IFM is reduced to the
    square lattice IFM.

\subsection{Proximity Graphs}
\label{ssec:graphtypes}
    A graph \(G(V,E)\) is a set of nodes \(V\) and edges \(E \subset V^{2}\).
    In the scope of this model, the nodes get identified with the sites of the
    lattice. If two nodes are connected by an edge, the sites are neighbors.\\
    The edges of \emph{proximity graphs} connect nodes which are by some metric near
    to each other. Hence they are suited to generalize problems defined on regular
    lattices with nearest neighbor relationships.
    To determine the distance we use the Euclidean
    metric on two-dimensional graphs, though in principle every metric in any
    dimension can be used.

    The \emph{Relative Neighborhood graph} (RNG) \cite{Toussaint1980} is
    a proximity graph. Two nodes \(i\) and \(j\) with distance $d_{ij}$
    will be connected, if no other node is in the \emph {lune}. The lune
    is defined as the intersection of two circles with radius \(r =
    d_{ij}\) and centers on \(i\) and \(j\). See also the hatched region
    in Fig.~\ref{fig:lunes}. An Example is given in Fig.~\ref{sfig:lunes:rng}.

    \begin{figure}[htbp]
        \centering
        \subfigure[]{
            \label{sfig:lunes:def}
            \includegraphics[width=0.14\textwidth]{images/GG_RNG_def}
        }
        \subfigure[]{
            \label{sfig:lunes:rng}
            \includegraphics[width=0.14\textwidth]{images/RNG/L12S03.pdf}
        }
        \subfigure[]{
            \label{sfig:lunes:gg}
            \includegraphics[width=0.14\textwidth]{images/GG/L12S03.pdf}
        }
        \caption[Gabriel and Relative Neighborhood Graph]
        {
            \subref{sfig:lunes:def} Lunes, which define whether an edge
                exist, of RNG (hatched region) and GG (cross hatched region).
                It is evident from this sketch that the GG is a supergraph
                of the RNG. If there is an edge in the RNG, the hatched
                region contains no node, then of course also the double
                hatched region contains no node and thus this edge appears
                also in the GG. So every egde of the RNG is also present
                in a GG on the same set of nodes.
            \subref{sfig:lunes:rng} Example of a RNG on periodic
                boundary conditions.
            \subref{sfig:lunes:gg} Example of a GG on
                periodic boundary conditions. Periodic nodes are dashed.
        }
        \label{fig:lunes}
    \end{figure}

    The \emph{Gabriel graph} (GG) \cite{Gabriel1969} is a subgraph
    of the Delaunay triangulation \cite{Delaunay1934,Katajainen}
    (DT), i.e.~for the same set of nodes \(V\) the edge set of the
    DT is a superset of the edge set of the GG. Two nodes \(i\) and
    \(j\) with distance \(d_{ij}\) will be connected with an edge,
    if a circle with its center on half way between \(i\) and \(j\)
    and radius \(r = \frac{d_{ij}}{2}\) contains no other nodes. See
    also the cross hatched region from Fig.~\ref{sfig:lunes:def}. An
    Example is given in Fig.~\ref {sfig:lunes:gg}. Note that the
    lune of the GG is completely enclosed in the lune of the RNG,
    therefore the RNG is a subgraph of the GG.

    To construct these graphs there is an algorithm\cite{RNGCell} with
    a time complexity of \(\mathcal{O}(N \log N)\) for nodes on a two
    dimensional plane.

    % TODO: alles hierdrunter
    %~ \subsubsection{Construction}
        %~ The simplest way to construct these graphs is to test for each
        %~ pair of nodes if any other node lies in
        %~ the lune of the pair. That is of complexity \(O (N^3)\), because
        %~ there are \(N(N-1)\) pairs and for each \((N-2)\) nodes to test. So
        %~ the running time of a straight forward implementation would be of
        %~ order \(O(N^3)\).
        %~ To reduce the running time one can basically use this simple method but only test
        %~ the connectedness for nodes which are near to the lune and abort if
        %~ one node inside the lune is found. To determine which nodes are
        %~ near the lune one can subdivide the area in \(M\) \emph{cells} of size
        %~ \(L_c \times L_c\) and save for each cell a list with nodes lying
        %~ inside like presented in Ref.\ \cite{RNGCell} for the RNG and sketched in Fig.\ \ref{fig:cell}.
        %~ For this research \(M = N\) is used.\\
        %~ \begin{figure}[htbp]
            %~ \centering
            %~ \subfigure[not Connected Nodes][]{
                %~ \label{sfig:cell:big}
                %~ \includegraphics[width=0.2\textwidth]{images/CellBig}
            %~ }
            %~ \subfigure[Connected Nodes][]{
                %~ \label{sfig:cell:small}
                %~ \includegraphics[width=0.2\textwidth]{images/CellSmall}
            %~ }
            %~ \caption[Sketch how the Cell Method works]
            %~ {
                %~ Sketch how the cell algorithm for the construction of the
                %~ RNG and GG works. Here with the lune definition of the GG.
                %~ The bounding box of the lune, which determines which cells have
                %~ to be tested, is marked with thick colored lines.
                %~ Left: shows that it is sufficient to find a
                %~ single node in an inner cell of the bounding box to discard
                %~ the edge between two distant nodes. The nodes inside the
                %~ other 8 marked cells do not have to be tested anymore.
                %~ Right: shows that only the nodes inside the
                %~ marked cell have to be tested, because there are no other nodes,
                %~ the edge can be drawn.
            %~ }
            %~ \label{fig:cell}
        %~ \end{figure}\\
        %~ Given the discretized cell structure it is just necessary to test the nodes in the cells which
        %~ resemble a rectangular bounding box of the lune. Most pairs will be
        %~ far away from each other and there will be one or more cells in the
        %~ middle of the bounding box, which are completely inside the lune.
        %~ Then only one node has to be tested from such a cell to discard an
        %~ edge between the nodes. On the other hand, nodes that will be connected
        %~ are near to each other so that only very few cells intersect the bounding
        %~ box of their lune and consequently only very few nodes have to be tested.\\
        %~ The complexity of this method should be between \(O(N^2)\) in the
        %~ best case, because for every pair at least one check has to be performed,
        %~ and \(O(N^3)\) in the worst case, e.g.\ all nodes are inside one
        %~ cell.\\
        %~ Note that even faster algorithms exsist for two dimensional RNG and GG, which require to first
        %~ generate a DT and use the fact, that the RNG and GG are subgraphs.
        %~ (For the construction of the DT for a given set of points one might
        %~ use existing software libraries, e.g.\ the \texttt{qHull}\cite{qhull} library.)\\
        %~ However, generation of the graphs is not time critical, because
        %~ the MC simulations take much longer. So only the simple cell algorithm
        %~ was used to gegnerate the graphs.\\